---
title: "Prosper Loan Exploratory Data Analysis"
author: Mustafa Olomi
date: 11 20, 2017
output: 
  html_document:
    toc: TRUE
    toc_depth: 3
    toc_float: TRUE
---


```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
library(ggplot2)
library(dplyr)
library(grid)
library(gridExtra)
library(reshape)
library(ggplot2)
library(ggthemes)
theme_set(theme_gdocs())
```
```{r}
ld <- read.csv('prosperLoanData.csv')
```

# **Introduction**: 
This is an exploratory analysis of the data from the peer to peer
lending company Prosper. The dataset contains over 100,000 borrowers with 84
variables. Prosper connects lenders, who are individual investors that can 
selectively invest their money in loans to borrowers based on a variety of 
criteria like credit score, prospers rating grade, profession, loan type & more.

In this analysis I aim to visualize and understand the characteristics of a 
good loan prospect and some of the qualities that make up bad loans, as well 
as Prosper's overall customer and loan profiles.

# Univariate Plots Section

In this section, I perform some preliminary exploration of the dataset

Firstly I take out the cancelled loans because there are only a few cancelled
loans and the loans were not initialized so we don't need them in our data

```{r}
#Taking out cancelled loans
ld <- subset(ld, LoanStatus != "Cancelled")
```

The avg credit score will give us a better datapoint to measure against

```{r}
#Averaging lower and upper credit score
ld$AvgCreditScore <- (ld$CreditScoreRangeUpper+ld$CreditScoreRangeLower) / 2
```


Plotting the distribution of avg credit scores. The middle line is the mean
Most traditional companies would want to know the credit score of the 
portfolio as it is a more universal measuring yardstick

```{r echo=FALSE, Univariate_Plots}
ggplot(ld, aes(AvgCreditScore)) +geom_histogram(color='black', 
                                                fill='light blue') +
  coord_cartesian(xlim = c(400,900)) +
  geom_vline(aes(xintercept = mean(AvgCreditScore, na.rm = T)),  
             color = 'red', linetype='dashed', size=1) +
  annotate("text", label = "Mean Credit Score 695.07", x = 600, y = 18000, 
           color = "red")
```

We can see the mean credit score for all loans is 695
This is considered "good" credit according to Experiean

300-579 is Very Poor (17% of people) may not be approved for credit at all
580-669 is Fair (20.2%) considered subprime
679-739 is Good (21.5%) 8% liklihood of serious delinquency
740-799 is Very Good (18.2%) likely to receive better than avg rates
800-850 is Exceptional(19.9%) at the top of the list for best rates w lenders

https://www.experian.com/blogs/ask-experian/credit-education/score-basics


Creating labels for each of the categories

```{r}
labs <- c('NA', 'Debt Cons', 'HomeImp', 'Biz', 'PerLoan', 'Student', 'Auto', 
          'Other', 'Baby', 'Boat', 'Cosmetic', 'EngRing', 'Green', 
          'HouseholdExp', 'LargePurch', 'Med/Dental', 'Motorcycle', 'RV', 
          'Taxes', 'Vaca', 'Wedding')

ld$Category <- factor(ld$ListingCategory..numeric., labels = labs)
```

Debt consolidation is the highest category with over 50% of the loans

```{r}
ld.cats <- ld %>%
    group_by(Category) %>%
    summarise(count = n()) %>%
    mutate(catpercs =count/sum(count))
    ld.cats %>% arrange(desc(catpercs))
    

ggplot(ld.cats, aes(x = reorder(Category, catpercs), y = catpercs*100)) +
  geom_bar(stat="identity", width = 0.7, fill = 'light blue') +
  theme_minimal(base_size = 14) +
  coord_flip()
```

```{r}
#Adding another column looking at just the date of the loan
ld$LoanDate <- as.Date(ld$LoanOriginationDate, "%Y-%m-%d")
```

In the documentation it says:
Pre  2009 prosper generated grades with the variable CreditGrade 
Post 2009 they used the grade ProsperRating Alpha 

Here we are creating a universal grading standard with all grades

```{r}
ld <- ld %>%
mutate(ProsperGradeAll = ifelse(ProsperRating..Alpha. != '', 
                                as.character(ProsperRating..Alpha.), 
                                as.character(CreditGrade)))
```

```{r}
ld <- subset(ld, ProsperGradeAll != 'NC' & ProsperGradeAll != '')
```

```{r}
prop.table(table(ld$ProsperGradeAll))
```


```{r}
ld$ProsperGradeAll <- factor(ld$ProsperGradeAll, 
                             levels = c('AA','A','B', 'C', 'D', 'E', 'HR'))
summary(ld$ProsperGradeAll)
```



```{r}
#First lets adjust the close date to be just the date 
#(since time is not in it anyway)
#First Recorded credit line also goes up to date level
ld$ClosedDate <- as.Date(ld$ClosedDate, "%Y-%m-%d")
ld$FirstRecordedCreditLine <- as.Date(ld$FirstRecordedCreditLine, "%Y-%m-%d")
```

We can see we have many categories, including 6 past due categories 
We could simplify this an better define the categories

```{r}
#Lets check what loan statuses we have
summary(ld$LoanStatus)
```

Based on the existing categories we can see that most of the status's are
centered around being complete, current or chargedoff/defaulted

According to Prosper's website: 
A borrower loan is charged-off when it reaches 121 days past due.

```{r}
ggplot(ld, aes(LoanStatus)) + geom_histogram(stat='count', fill = 'light blue') +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

Lets check to see where we should categorize the final payment loans 
Looking at their % funded
Because they are 99.65% funded we'll categorize them as 'Complete'

```{r}
summary(subset(ld, LoanStatus=='FinalPaymentInProgress')$PercentFunded)
```

Because the data is centered around complete, current and chargedoff/defaulted
I am regrouping the data below into those categories

Summarizing all past dues as one category so we can easily see them together
Summarizing charged off & defaulted as one category because they are similar

With this we can see the loans as a narrower category range and better view the
loans and what standing they are in

```{r}
#When I used the case_when, it does this as a character instead of a vector
#So changing to vector after case_when
ld <- mutate(ld, Status = case_when (LoanStatus == 'Current' ~ 'Current',
                                     LoanStatus == 'Completed' ~ 'Completed',
                                     LoanStatus == 'FinalPaymentInProgress' & 
                                       PercentFunded >= 0.95 ~ 'Completed',
                                     LoanStatus == 'FinalPaymentInProgress' & 
                                       PercentFunded < 0.95 ~ 'Past Due',
                                     LoanStatus %in% c('Chargedoff', 'Defaulted') 
                                     ~ 'Defaulted',
                                     LoanStatus %in% c('Past Due (1-15 days)', 
                                                       'Past Due (16-30 days)', 
                                                       'Past Due (31-60 days)', 
                                                       'Past Due (61-90 days)', 
                                                       'Past Due (91-120 days)', 
                                                       'Past Due (>120 days)') 
                                     ~ 'Past Due'))

ld$Status <- factor(ld$Status, levels = c('Completed', 'Current', 
                                          'Past Due', 'Defaulted'))
summary(ld$Status)
```


This is the distribution of completed, current, past due and defaulted loans

```{r}
ggplot(ld, aes(Status)) + geom_histogram(stat='count', fill = 'light blue') +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

This is the same distribution but is a percentage distribution

```{r}
ggplot(ld, aes(x=Status)) +
  geom_histogram(data = ld, stat = 'count', 
                 aes(y = ..count.. / sum(..count..)), fill = 'light blue') + 
  scale_y_continuous(labels = scales::percent)
```


```{r}
summary(ld$ProsperGradeAll)
```

Interesting to see that it does look like a normal distribution curve
Even though their credit ratings are not as normal

```{r}
ggplot(ld, aes(ProsperGradeAll)) + geom_histogram(stat = 'count', 
                                                  fill = 'light blue') +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

#Bivariate Analysis

Now looking at the interest rate among the 4 main categories we set 
The top 75 percentile of completed loans were 25.11%
It's also interesting to see that that is near the median of bad loans
Past due and defaulted loans have a median just under 25.11% 
So this is a good indicator of when a loan may be too expensive or too
high risk for clients.

```{r}
ggplot(ld, aes(Status, BorrowerRate)) +geom_boxplot(fill = 'light blue') +
  geom_hline(yintercept = quantile(subset(ld, Status=='Completed')$BorrowerRate, 
                                   probs = 0.75, na.rm = TRUE), color = 'red',
             linetype='dashed') +
    annotate("text", label = " 25.11%", 
           x = 0.8, y = 0.27, color = "red", size=3.5)
```

```{r}
quantile(subset(ld, Status=='Completed')$BorrowerRate, 
                                   probs = 0.75, na.rm = TRUE)
```

This shows the distribution of prosper grades and interest rates
We can see that 24.2 is the median rate for D grade loans
Interesting that this is close to the median of past due/defaulted loans
23.85 is the median for defaulted vs 24.19 for D grade loans

```{r}
ggplot(ld, aes(ProsperGradeAll, BorrowerRate)) + 
  geom_boxplot(fill = 'light blue') +
  theme(text = element_text(size = 12), 
        axis.text.x = element_text(angle = 35, hjust = 1)) +
  geom_hline(yintercept=median(subset(ld, ProsperGradeAll=='D')$BorrowerRate)) +
    #geom_hline(yintercept=median(ld$StatedMonthlyIncome)) +
  scale_fill_brewer(palette="Pastel1") +
  annotate("text", label = "24.2%", 
           x = 'D', y = 0.28, color = "blue", size=3.5) +
    geom_hline(yintercept = median(subset(ld, Status=='Defaulted')$BorrowerRate)
               , color = 'red',
             linetype='dashed')
```


```{r}
median(subset(ld, ProsperGradeAll=='D')$BorrowerRate)
```

```{r}
median(subset(ld, Status=='Defaulted')$BorrowerRate)
```

Unemployed borrowers had a median rate of 26%, higher than loans to employed

```{r}
ggplot(ld, aes(EmploymentStatus, BorrowerRate),) + 
  geom_boxplot(fill = 'light blue') +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_hline(aes(yintercept=median(subset(ld, EmploymentStatus=='Not employed')
                                   $BorrowerRate)),
             color='red', linetype='dashed') +
  annotate("text", label = "Median Borrower Rate for Not Employed = 26%", 
           x = 'Not employed', y = 0.37, color = "blue", size=3.5)
```

```{r}
median(subset(ld, EmploymentStatus=='Not employed')$BorrowerRate)
```


```{r}
median(subset(ld,  Status=='Defaulted')$BorrowerRate)
```

The most interesting relationship I saw was the fact that bad loans tended
to be around the 25% interest rate level, thats the median for defaulted 
loans, and around the median for D grade loans and for unemployed borrowers

Whatever the case, a borrower that is willing to pay close to 25% is usually
a sign of financial irresponsibility that leads to bad performing loans


# Multivariate Plots Section


When examining the income rates of employed individuals and their rates
We can see that a good floor for income level is $3000
This is the median income for defaulted loans in the HR category
It is also the bottom 25% of most almost every other grades defaulted loan


```{r}
ggplot(subset(ld, EmploymentStatus != 'Not employed'), 
       aes(ProsperGradeAll, StatedMonthlyIncome, fill=Status)) + 
  geom_boxplot() +
  theme(text = element_text(size = 12), 
        axis.text.x = element_text(angle = 35, hjust = 1)) +
  coord_cartesian(ylim = c(0,10000)) +
  geom_hline(yintercept=median(subset(ld, Status == 'Defaulted' & 
                                        ProsperGradeAll == 'HR' &
                                        EmploymentStatus != 'Not employed')
                                $StatedMonthlyIncome), linetype = 'dashed') +
  scale_fill_brewer(palette="Pastel1")  +
  annotate("text", label = "Median Income Defaulted HR Employed = 3000", 
           x = 'E', y = 1800, color = "blue", size=3.5)

```

```{r}
median(subset(ld, Status == 'Defaulted' & ProsperGradeAll == 'HR' &
                                        EmploymentStatus != 'Not employed')
       $StatedMonthlyIncome)
```



```{r}
median(subset(ld, ProsperGradeAll=='C' & IncomeVerifiable=='False')
       $StatedMonthlyIncome)
```


It looks like the Prosper grades also shifted like the way credit scores did
Just because the above graph resembles a normal distribution, doesn't mean
that it is a normal distribtion, credit standards changed after 2009
We will explore this in the bivariate plot section also

```{r}
ggplot(ld, aes(x=ProsperGradeAll)) +
  geom_histogram(data = subset(ld, ClosedDate= '2009-01-01'), stat = 'count', 
                 aes(y = ..count.. / sum(..count..)), fill = 'blue', 
                 alpha = 0.4, width = 0.8) + 
  geom_histogram(data = subset(ld, ClosedDate< '2009-01-01'), stat = 'count', 
                 aes(y = ..count.. / sum(..count..)), fill = 'green', 
                 alpha = 0.5, width = 0.6)+
  scale_y_continuous(labels = scales::percent)
```


I am looking at the data before 2009 and after 2009
Prosper went through some changes because of the credit crisis and we can
see how they changed their lending criteria before and after 2009

```{r}
ggplot(ld, aes(x=AvgCreditScore)) +
  geom_histogram(data = subset(ld, ClosedDate= '2009-01-01'), stat = 'count', 
                 aes(y = ..count.. / sum(..count..)), fill = 'blue', 
                 alpha = 0.4, width = 17) + 
  geom_histogram(data = subset(ld, ClosedDate< '2009-01-01'), stat = 'count', 
                 aes(y = ..count.. / sum(..count..)), fill = 'green', 
                 alpha = 0.5, width = 8)+
  scale_y_continuous(labels = scales::percent) +
    coord_cartesian(xlim = c(400, 900)) + 
  geom_vline(data = subset(ld, ClosedDate= '2009-01-01'), 
             aes(xintercept = mean(AvgCreditScore, na.rm = T)), 
             color = 'blue', linetype='dashed', size=0.5) +
  geom_vline(data = subset(ld, ClosedDate < '2009-01-01'), 
             aes(xintercept = mean(AvgCreditScore, na.rm = T)), 
             color = 'dark green', linetype='dashed', size=1) +
  annotate("text", label = "Mean Avg Credit Pre 2009 = 643.61", 
           x = 540, y = 0.105, color = "darkgreen", size=3.5) +
  annotate("text", label = "690.61 = Mean Avg Credit Post 2009", 
           x = 825, y = 0.105, color = "blue", size=3.5)
```


We can see that it looked like a fairly normal distribution prior to 2009
After 2009 the cut down on loans under 600 credit score
Their average credit score also increased from 653.59 to 708.91

```{r}
mean(subset(ld, ClosedDate = '2009-01-01')$AvgCreditScore, na.rm = T)
mean(subset(ld, ClosedDate < '2009-01-01')$AvgCreditScore, na.rm = T)
```


Because debt consolidation is the most popular loan category it is worth
exploring this category more and seeing what trends make successful loans

Below we can see the debt to income of Completed vs Defualted loans
Defaulted loans have 24% higher mean debt to income ratio 25% vs 31%

```{r}
ggplot(subset(ld, Category == 'Debt Cons' & Status %in% 
                c('Completed', 'Defaulted')), 
       aes(DebtToIncomeRatio, fill = Status)) +
  geom_bar(stat = 'count') +
  coord_cartesian(xlim = c(0,0.75)) +
  scale_fill_brewer(palette="Pastel1") +
  geom_vline(aes(xintercept=mean(subset(ld, Category == 'Debt Cons' 
                                          & Status == 'Completed')$
                                   DebtToIncomeRatio, na.rm = TRUE)), 
             color='red', linetype='dashed') +
  annotate("text", label = "Mean Debt/Income 0.25", x = 0.16, y = 200, 
           color = "red", size=3.5) +
  geom_vline(aes(xintercept=mean(subset(ld, Category == 'Debt Cons' 
                                          & Status == 'Defaulted')$
                                   DebtToIncomeRatio, na.rm = TRUE)), 
             color='blue', linetype='dashed') +
  annotate("text", label = "Mean Debt/Income 0.31", x = 0.41, y = 120, 
           color = "blue", size=3.5)
```


```{r}
mean(subset(ld, Category == 'Debt Cons' 
                                          & Status == 'Completed')
     $DebtToIncomeRatio, na.rm = TRUE)

mean(subset(ld, Category == 'Debt Cons' 
                                          & Status == 'Defaulted')
     $DebtToIncomeRatio, na.rm = TRUE)
```

Looking at the bank card utilizatin rate at completed and defaulte loans
We can see that unsuccessful loans had 22% higher bank card utilization
50% vs 61%

```{r}
ggplot() +
  geom_bar(data = subset(ld, Category == 'Debt Cons'), aes(BankcardUtilization), 
           stat = 'count', fill = 'light blue') +
  geom_bar(data = subset(ld, Category != 'Debt Cons'), aes(BankcardUtilization), 
           stat = 'count', fill = 'gray') +
      geom_vline(aes(xintercept=mean(subset(ld, Category == 'Debt Cons')$
                                   BankcardUtilization, na.rm = TRUE)), 
             color='blue', linetype='dashed') +
      geom_vline(aes(xintercept=mean(subset(ld, Category != 'Debt Cons')$
                                   BankcardUtilization, na.rm = TRUE)),
                 color='black', linetype='dashed') +
  coord_cartesian(xlim = c(0,1.5), ylim = c(0,1500)) +
    annotate("text", label = "Mean Bankcard Util = 0.50", x = 0.29, y = 240, 
           color = "black", size=3.5) +
    annotate("text", label = "Mean Bankcard Util = 0.61", x = 0.82, y = 100, 
           color = "blue", size=3.5) +
  scale_fill_brewer(palette="Pastel1")
```

```{r}
mean(subset(ld, Category == 'Debt Cons')$BankcardUtilization, na.rm = TRUE)
mean(subset(ld, Category != 'Debt Cons')$BankcardUtilization, na.rm = TRUE)
```


The loan quarter is formatted so that the Q is written first then the year
Thus below we are transforming it so the year is frst then the Q


```{r}
ld <- mutate(ld, LoanQtr = paste(substring(LoanOriginationQuarter, 4,8), 
                                 substring(LoanOriginationQuarter, 1,2)))
```

I visualized the loan payments by quarter by grade so we can see where their 
cash flows are coming from on a rolling quarter to quarter basis. We can see 
that C grade loans have grown more popular ad received more funding recently 
while B and A loans have remained consistent, and the number of AA loan 
flows droppedd after 2009 perhaps because the credit market recovered and 
good borrowers have refinanced or found other funding options.

```{r}
ggplot(ld, aes(LoanQtr, LP_CustomerPayments/1000000, fill = ProsperGradeAll)) +
  geom_bar(stat = 'identity') +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 10)) +
  scale_fill_brewer(palette="Pastel1")
```

Below we can see the loan amounts (in millions) based on grade so we can see
the quality of the loans originated as well as the amount of the loan.
We see that A, B, & C loans all grew in terms of loans originated near 2013

```{r}
ggplot(ld, aes(LoanQtr, LoanOriginalAmount/1000000, fill = ProsperGradeAll)) +
  geom_bar(stat='identity') +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, size=10)) +
  scale_fill_brewer(palette="Pastel1")
```

Now we can look at each loan category and see if they are in good standing
We see the distribution in millions of how many million are in each status

```{r}
ggplot(ld, aes(ProsperGradeAll, LoanOriginalAmount/1000000, fill = Status)) +
  geom_bar(stat='identity') +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, size=11)) +
  scale_fill_brewer(palette="Pastel1")
```

Below we can see what percentage of each status is complete, past due etc.
It looks like the largest incremental percentage jump is between C and D
This combined with our earlier analysis that showed D loans as having
a median interest rate similar to the defaulted status tells us C is 
probably the lowest category of loan that is a good investment.

```{r}
ld2 <- ld %>% 
  group_by(ProsperGradeAll,Status) %>% 
  summarise(count=n()) %>% 
  mutate(perc=count/sum(count))


ggplot(ld2, aes(x = factor(ProsperGradeAll), y = perc*100, fill = factor(Status))) +
  geom_bar(stat="identity", width = 0.7) +
  theme_minimal(base_size = 14) +
  scale_fill_brewer(palette="Pastel1")
```

Below we see the different occupations of loan holders and what percentage
of those loans are in good or bad standing

Students at technical schools have the worst performing loans of all
Judges, Psycholigst and Dentists have the lowest default ratings

Further analysis could be done with students where we take into account
Schools they attend, graduation rates, grades, and job placement data

```{r}
ld3 <- ld %>% 
  group_by(Occupation,Status) %>% 
  summarise(count=n()) %>% 
  mutate(perc=count/sum(count))

ggplot(ld3, aes(factor(Occupation), y = perc*100, fill = Status)) + 
  geom_bar(stat='identity') +
  theme(axis.text.y = element_text(size=7)) + 
  coord_flip() +
  scale_fill_brewer(palette="Pastel1")
```

Now you can see the loan based on loan reason and the current standing
N/A and personal loans are the worst performing with students also
Continue to perform badly and have high defaults.

RV loans perform much better than all other categories perhaps because
RV customers are usually retired and older and trying to finance their
luxury purchase longer term.


```{r}
ld4 <- ld %>% 
  group_by(Category,Status) %>% 
  summarise(count=n()) %>% 
  mutate(perc=count/sum(count))

ggplot(ld4, aes(factor(Category), y = perc*100, fill = Status)) + 
  geom_bar(stat='identity') +
  coord_flip()+
  theme(axis.text.y = element_text(size=12)) +
  scale_fill_brewer(palette="Pastel1")
```


# Final Plots and Summary

First I think it's important to see Prosper's credit rating distribution
This plot shows both the pre credit crisis and post credit crisis data
It's good to see how Prosper adjusted their standards afterwards

Prosper's adjustments show us that they initially took on many loans of
all different credit categories to gain marketshare but after 2009
they really tightened their lending policy away from bad credit scores

```{r echo=FALSE, Plot_One}

ggplot(ld, aes(x=AvgCreditScore), xlab('Credit Score')) +
  geom_histogram(data = subset(ld, ClosedDate= '2009-01-01'), stat = 'count', 
                 aes(y = ..count.. / sum(..count..)), fill = 'blue', 
                 alpha = 0.4, width = 17) + 
  geom_histogram(data = subset(ld, ClosedDate< '2009-01-01'), stat = 'count', 
                 aes(y = ..count.. / sum(..count..)), fill = 'green', 
                 alpha = 0.5, width = 8)+
  scale_y_continuous(labels = scales::percent) +
  ggtitle("Credit Score Percentage Distribution") +
  coord_cartesian(xlim = c(400, 900)) + 
  geom_vline(data = subset(ld, ClosedDate= '2009-01-01'), 
             aes(xintercept = mean(AvgCreditScore, na.rm = T)), 
             color = 'blue', linetype='dashed', size=0.5) +
  geom_vline(data = subset(ld, ClosedDate < '2009-01-01'), 
             aes(xintercept = mean(AvgCreditScore, na.rm = T)), 
             color = 'dark green', linetype='dashed', size=1) +
  annotate("text", label = "Mean Avg Credit Pre 2009 = 653.59", 
           x = 540, y = 0.105, color = "darkgreen", size=3.5) +
  annotate("text", label = "708.91 = Mean Avg Credit Post 2009", 
           x = 825, y = 0.105, color = "blue", size=3.5) +
  labs(y = 'Percent Distribution', x = 'Avg Credit Score')
```


This chart shows how much in loans prosper was generating and by what grade
Its good to see that they figured out what kind of loans are most successful
Then after 2010 they ramped up their loans and we can see how C grade loans
grew the most, this is consistent with our analysis in seeing the C is the
lowest grade of loan that does not have a large incremental rise in defaults

```{r echo=FALSE, Plot_Two}
ggplot(ld, aes(LoanQtr, LoanOriginalAmount/1000000, fill = ProsperGradeAll)) +
  geom_bar(stat='identity') +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, size=10)) +
  scale_fill_brewer(palette="Pastel1") +
  ggtitle("Loans Originated in Millions by Quarter and Grade") +
  labs(y = 'Loan Origination amount in Millions', x = 'Quarter')
```

Finally here we can see the performance of each loan category
Its interesting to see different categories and how their loans perform
NA and Personal loans perform the worst perhaps because the reason for the
loan is not specified and because borrowers that do not have a specific
reason for the loan. 

Luxury items like RV, Motorcycle, and Boats perform the best because people
would not buy these items unless they could afford the luxury

Wedding, engagement rings, etc are all also times that require people to be
a lot more responsible so these loans look like they are more successful

```{r echo=FALSE, Plot_Three}
ld4 <- ld %>% 
  group_by(Category,Status) %>% 
  summarise(count=n()) %>% 
  mutate(perc=count/sum(count))

ggplot(ld4, aes(factor(Category), y = perc*100, fill = Status)) + 
  geom_bar(stat='identity') +
  coord_flip()+
  theme(axis.text.y = element_text(size=12)) +
  scale_fill_brewer(palette="Pastel1") +
  ggtitle('Status Percentages of Loan by Category') +
  labs(y = 'Percent', x = 'Loan Category')
```

#Reflection

Overall I learned a lot about the different criteria's that could be used
to create a good portfolio of loans that avoids parameters that usually
lead to defaults.

I would look for loans that are below the 25% rate and C grade or above
Also people who are employed and make $3000 per month and certain categories

For debt consolidation loans I would look at credit card utilization rates
and debt to income rate that are closer to that of complete or current loans

I was surprised to see that loans for luxury goods had such high completions 
Some of the struggles I had were that defaulted and completed loans had many
of the same characteristics, and if I were to go into more detail I would
do a more thorough analysis within each type of loan

I think having more unique data points like what types of schools students
were applying from and the students stats like their grades and majors
would be helpful in finding successful loans

I think there is no reason to give N/A loans, and personal loans need to 
have more of a description of what the funds are being used for

To take this analysis further I would backtest the parameters I found for 
successful loans and come up with a factor analysis to see how the different
returns are currently being calculated and look for possible arbitrage
opportunities with underpriced loans. I would create a model portfolio
and see show the risk level and what the overall percentage return would be

It was great to see that we could get such detailed analysis of the loans
I would think that many investors just pick loans based on categories they 
think are successful. I would not have lent to luxry items like boats
 because I live with a high level of financial responsibility and know
that boats are a bad investment, but the loan data suggests otherwise.

I think the skills I learned here will really shape how I look at data and 
decisions in the future. Scraping the data with python for any decision
and then exploring it with R will lead me to make much better decisions
in my personal and professional life and there is no excuse to make an 
uninformed and non data-driven decision in the future.
